Safe Transferable Regions

### 2018-06-05 paper/abstract.tex
Format changes.

There is an increasing interest in alternative memory management
schemes that seek to combine the convenience of garbage collection
and the performance of manual memory management in a single language
framework.  Unfortunately, ensuring safety in presence of manual
memory management remains as great a challenge as ever. In this paper,
we present a C\#-like object-oriented language called \name that uses
a combination of region type system and lightweight runtime checks to
enforce safety in presence of user-managed memory regions called
\emph{transferable regions}. Unsafe transferable regions have been
previously used to contain the latency due to unbounded GC pauses. Our
approach shows that it is possible to restore safety without
compromising on the benefits of transferable regions.  We prove the
type safety of \name in a formal framework that includes its
C\#-inspired features, such as higher-order functions and generics. We
complement our type system with a type inference algorithm, which
eliminates the need for programmers to write region annotations on
types. The inference algorithm has been proven sound and relatively
complete. We describe a prototype implementation of the inference
algorithm, and our experience of using it to enforce memory safety in
dataflow programs.
% %
% In this paper, we present a language and a memory management scheme based
% on user-managed memory regions (called \emph{transferable regions})
% that allow programmers to exploit knowledge of data structures' lifetimes
% to improve performance.

>>>>

There is an increasing interest in alternative memory management
schemes that seek to combine the convenience of garbage collection and
the performance of manual memory management in a single language
framework.  Unfortunately, ensuring safety in presence of manual
memory management remains as great a challenge as ever. In this paper,
we present a C\#-like object-oriented language called {\sc
Broom}\xspace that uses a combination of region type system and
lightweight runtime checks to enforce safety in presence of
user-managed memory regions called \emph{transferable regions}. Unsafe
transferable regions have been previously used to contain the latency
due to unbounded GC pauses. Our approach shows that it is possible to
restore safety without compromising on the benefits of transferable
regions.  We prove the type safety of {\sc Broom}\xspace in a formal
framework that includes its C\#-inspired features, such as
higher-order functions and generics. We complement our type system
with a type inference algorithm, which eliminates the need for
programmers to write region annotations on types. The inference
algorithm has been proven sound and relatively complete. We describe a
prototype implementation of the inference algorithm, and our
experience of using it to enforce memory safety in dataflow programs.
% %
% In this paper, we present a language and a memory management scheme based
% on user-managed memory regions (called \emph{transferable regions})
% that allow programmers to exploit knowledge of data structures' lifetimes
% to improve performance.

### 2018-06-05 paper/broom.tex
Edits to appendix. CR04 added as related work.
Ecoop18-related changes and the rebuttal.
Format changes.

\documentclass[a4paper,UKenglish]{lipics-v2016}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
% for section-numbered lemmas etc., use "numberwithinsect"

@@

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
\author[1]{Gowthan Kaki}
\author[2]{G. Ramalingam}
\author[3]{Kapil Vaswani}
\author[4]{Dimitrios Vytiniotis}
\affil[1]{Purdue University}
\affil[2]{Microsoft Research}
\affil[3]{Microsoft Research}
\affil[4]{Microsoft Research}
\authorrunning{G. Kaki, G. Ramalingam, K. Vaswani, and D. Vytiniotis} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{Gowtham Kaki, G. Ramalingam, Kapil Vaswani, and Dimitrios Vytiniotis}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

% \subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
\keywords{Memory management, regions, type system, type inference}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{ECOOP 2017}
\EventAcronym{ECOOP}
\EventYear{2017}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\documentclass[a4paper,UKenglish]{lipics-v2018}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
% for section-numbered lemmas etc., use "numberwithinsect"

@@

\author{Gowtham Kaki}
       {Purdue University, USA\footnote{Work done during an internship at
Microsoft Research, India.}}
       {gkaki@purdue.edu}
       {}
       {}%mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
\author{G. Ramalingam}
       {Microsoft Research, India}
       {grama@microsoft.com}
       {}
       {}%mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.

\authorrunning{G. Kaki and G. Ramalingam}%mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Gowtham Kaki and G. Ramalingam}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\subjclass{
	\ccsdesc[500]{Software and its engineering~Allocation / deallocation strategies};\xspace
	\ccsdesc[500]{Software and its engineering~Object oriented languages};\xspace
	\ccsdesc[500]{Software and its engineering~Data flow languages};\xspace
	\ccsdesc[300]{Software and its engineering~Software verification and validation};\xspace
}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
\keywords{Memory Safety, Formal Methods, Type System, Type Inference,
Regions, Featherweight Java}% mandatory: Please provide 1-5 keywords

% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgements{We are grateful to Kapil Vaswani, Dimitrios
Vytiniotis, Michael Isard, and Steven Hand for their valuable inputs
during the initial stages of this project. We would also like to thank
the anonymous reviewers for their careful scrutiny and insightful
comments that helped improve the quality of this paper.}%optional

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{Todd Millstein}
\EventNoEds{1}
\EventLongTitle{32nd European Conference on Object-Oriented Programming (ECOOP 2018)}
\EventShortTitle{ECOOP 2018}
\EventAcronym{ECOOP}
\EventYear{2018}
\EventDate{July 16--21, 2018}
\EventLocation{Amsterdam, Netherlands}
\EventLogo{}
\SeriesVolume{109}
\ArticleNo{11}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
